---
title: "Data Science Challenge"
author:
- name: Gopalakrishnan Kalarikovilagam Subramanian
date: "July 9, 2019"
output: 
  prettydoc::html_pretty:
    theme: architect
---
  
## Introduction

In this data challenge there are 28 variables provided to us. They are:
  
  ***Key variables used***:
  
* duration	Duration in minutes

* director_name	Name of the Director of the Movie

* director_facebook_likes	Number of likes of the Director on his Facebook Page

* actor_1_name	Primary actor starring in the movie

* actor_1_facebook_likes	Number of likes of the Actor_1 on his/her Facebook Page

* cast_total_facebook_likes	Total number of facebook likes of the entire cast of the movie

* movie_facebook_likes	Number of Facebook likes in the movie page

* genres	Film categorization like Animation, Comedy, Romance, Horror, Sci-Fi

* title_year	The year in which the movie is released (1916:2016)

* content_rating	Content rating of the movie

* gross	Gross earnings of the movie in Dollars

* budget	Budget of the movie in Dollars

The objective is to find the correct recipe for creating a blockbuster movie.
## Packages Required

The following packages are being used for our analysis:
  
```{r message = FALSE}
rm(list=ls())
library(readxl)
library(tidyverse)
library(dplyr)
require(gdata)
```

## Data Importing 

The following code is used to import the dataset in r after placing them in the 
working directory. 

```{r message = FALSE, results='hide', eval = TRUE}
#import data
Movie = read.xls ("movie_metadata.xls", sheet = 1, header = TRUE)

```

## Data Cleaning

The data set is cleaned to remove duplicates.
```{r message = FALSE, results='hide', eval = TRUE}
# To remove duplicates from the data

Movie <- Movie[!duplicated(Movie[]),] 

# To create the  profit column
Movie$profit <- Movie$gross - Movie$budget

```
After data collection it is tidied exploration is done on the data.

## Data Exploration

The summary of the data set is performed. It is overwhelmingly color movies.
The shortest film is 7 min and the longest is 511. 50% of the movies are in the 
93 min to 118 min range. Actor 1 has on an averge more facebook likes than
other crew members.

The lowest amount grossed by a movie is 162 and the highest is more than 760 million.
Drama is the highest genre followed by comedy and comedy/drama.

It is found that most of the movies are in english language, followed by french and spanish. USA produces most movies,
followed by UK and France. Most of the movies are R rated, followed by PG-13.
The lowest budget movie is 218 which seems very small. 75% of the movies in the
data set are released after 1999 .The score of the movie range from 1.6 to 9.5.
The aspect ratio of the movies range from 1.18 to 16.

```{r message = FALSE, results='hide', eval = TRUE}
summary(Movie)

```

The plot of top five directors as a function of avaerage profit is done. The 
directors are filtered on who have more than 5 movies. This gives more 
reliability on the part of the directors. The top five reliable directors are
James Cameroon, Jon Favreau, Steven Spielberg, Christopher Nolan and Jay Roach.

A similar plot is done for first actors and these are the most protitable actors:
  
  Robert Pattison, Bryce Dallas Howard, Steve Carell, Harrison Ford and Jennifer  
Lawrence.  

The movies are classified into foir categories : very small if it is less than an
hour. Small is it is less than 2 hours, Medium if less than 3 hours and Long if
more than 3 hours. A plot of average profit of the movie based on this 
classification is carried out. It shows that very-short movies are not at all 
profitable. The most profitable are the medium range movies followed by long range 
and short movies.

A plot of average profit versus genre of the movie is carried out. The Sci-Fi|Family
movies are the most profitable, followed by Adventure|Biography|Drama|Family|Musical.

The actor_1 with top facebook likes are explored. The top five are Darcy Donavan,
Matthew Ziff, Krista Allen, Andrew Fiscella and Jimmy Bennet, Michael Joiner, 
Robin Williams, Craig Stark, Muse Watson and Tim Holmes. 

Since Facebook was very famous after 2006 a plot of top 10 profitable actors 
based on movies released aster 2006 is carried out. None of the actors feature in
the actors with top facebook likes. So the number of facebook likes does not
seem to have a bearing on the profitability of the actors.

A plot of profitability and movie score is carried out. The average profit varies
exponentially with movie score. This is an importane parameter which determines
profitability of a movie.

A plot of profitability and total cast facebook likes and movies facebook likes
is plotted. There does not seem to be a relationship between the facebook likes
and profitability.

```{r message = FALSE, results='hide', eval = TRUE, warnings = FALSE}
# Top 5 profitable directors
Movie %>% group_by(director_name) %>% summarise(count = n(), avg_profit = mean(profit, na.rm = TRUE)) %>%
  filter(count>5) %>% arrange(-avg_profit) %>% slice(1:5) %>%
  ggplot(aes(x = director_name, y = avg_profit)) + geom_bar(stat="identity")

# Top five profitable first actor
Movie %>% group_by(actor_1_name) %>%
  summarise(count = n(), avg_profit = mean(profit, na.rm = TRUE)) %>%
  filter(count>5) %>% arrange(-avg_profit) %>% slice(1:5) %>%
  ggplot(aes(x = actor_1_name, y = avg_profit)) + geom_bar(stat="identity")

# Classifying movies into different categories and seeing the variation of profit
Movie %>% 
  mutate(dur = ifelse(duration<60, 'very short', ifelse(duration<120, 'short', ifelse(duration<180, 'medium', 'long')))) %>% 
  group_by(dur) %>% filter(all(!is.na(dur))) %>%
  summarise(avg_profit = mean(profit, na.rm = TRUE)) %>%
  ggplot(aes(x = dur, y = avg_profit)) + geom_bar(stat="identity")

# Classifying movies into different genres and seeing the variation of profit
Movie %>% 
  group_by(genres) %>% 
  summarise(avg_profit = mean(profit, na.rm = TRUE)) %>% arrange(-avg_profit) %>% 
  slice(1:10) %>%
  ggplot(aes(x = genres, y = avg_profit)) + geom_bar(stat="identity") +
  coord_flip()

Movie %>% 
  group_by(actor_1_name) %>% filter(actor_1_facebook_likes>11000) %>% 
  summarise(avg_profit = mean(profit, na.rm = TRUE)) %>% arrange(-avg_profit) %>%
  slice(1:5) %>%
  ggplot(aes(x = actor_1_name, y = avg_profit)) + geom_bar(stat="identity")

Movie %>% group_by(actor_1_name) %>% 
  summarise(avg_likes = mean(actor_1_facebook_likes, na.rm = TRUE)) %>%
  arrange(desc(avg_likes)) %>%
  slice(1:10) %>% subset(select = c(actor_1_name, avg_likes))

# Classification of profit based on actor_1 for movies after 2006 and having 5 or more movies

Movie %>% 
  filter(title_year > 2006) %>% group_by(actor_1_name) %>%
  summarise(count = n(), avg_profit = mean(profit, na.rm = TRUE)) %>% 
  filter(count>5) %>%
  arrange(-avg_profit) %>%
  slice(1:10) %>%
  ggplot(aes(x = actor_1_name, y = avg_profit)) + geom_bar(stat="identity")+coord_flip()

# Profitability and movie score

Movie %>%  group_by(movie_score) %>% 
  summarise(avg_profit = mean(profit, na.rm = TRUE)) %>%
  filter(avg_profit > 0) %>% 
  ggplot(aes(x = movie_score, y = avg_profit))+geom_point()

# Profitability and movie_facebook_likes
Movie %>%  filter(title_year > 2006) %>% group_by(movie_facebook_likes) %>% 
  summarise(avg_profit = mean(profit, na.rm = TRUE)) %>%
  filter(avg_profit > 0) %>% 
  ggplot(aes(x = movie_facebook_likes, y = avg_profit))+geom_point()

# Profitability and cast_total_facebook_likes

Movie %>%  filter(title_year > 2006) %>% group_by(cast_total_facebook_likes) %>% 
  summarise(avg_profit = mean(profit, na.rm = TRUE)) %>%
  filter(avg_profit > 0) %>% 
  ggplot(aes(x = cast_total_facebook_likes, y = avg_profit))+geom_point()

```

## Summary

The work can be summarized as follows:
  
1. James Cameroon, Jon Favreau, Steven Spielberg, Christopher Nolan and Jay Roach 
are top 5 profitable directors.

2. Robert Pattison, Bryce Dallas Howard, Steve Carell, Harrison Ford and 
Jennifer Lawrence are the top 5 profitable first actors.

3. Making a medium duration (1hr to 2 hr) movie is highly profitable.

4. Family|Sci-Fi is the most profitable genre.

5. M rated movies are the most profitable.

6. There is no relationship between Facebook likes and profitability. 
So cost spent on these platforms can be avoided.

7. Movie score is the most important factor determining the profitability of a 
movie.

8. If a movie ticks all the above boxes, it is sure to be profitable.





